---
title: "Stat 536 HW5"
author: "Dongyang Wang"
date: "2022-11-10"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Question 1

```{r}
df1 = c(2037,1757,958,218)
df1 = array(df1, c(2,2))

indep_mod1 <- loglin(df1,margin=list(1,2),fit=TRUE,param=TRUE)
```
The independence model does not seem to fit well since the prediction is a bit off. We fit the saturated model instead. 

```{r}
sat_mod1 <- loglin(df1,margin=list(c(1,2)),fit=TRUE,param=TRUE)
```


## Question 2

$\log \frac{P(X_1 =Yes | X_2 = k)}{P(X_1 =No | X_2 = k)} = u_{1(1)} - u_{1(2)} = 0.2081879 + 0.2081879 =0.4163758$. The interpretation is that the odds of yes over no are exp(0.4163758) = 1.516456.

```{r}
exp(0.4163758)
```


## Question 3

Under the independence model, we reject the null.and there is indeed a difference. The alternative hypotheses include: BB is different from Bb only; BB is different from bb only; BB is different from Bb and bb.

```{r}
1- pchisq(indep_mod1$lrt, indep_mod1$df)
```


## Question 1

The asymptotic result is calculated by subtracting the lrt's and df's. 
```{r}
df2 = c(2037,1757,631, 18, 327, 200)
df2 = array(df2, c(2,3))

indep_mod2 <- loglin(df2,margin=list(1,2),fit=TRUE,param=TRUE)
sat_mod2 <- loglin(df2,margin=list(c(1,2)),fit=TRUE,param=TRUE)

1-pchisq(indep_mod2$lrt-sat_mod2$lrt,indep_mod2$df-sat_mod2$df)
```


## Question 2

```{r}
1-pchisq(indep_mod2$lrt,indep_mod2$df)
```


## Question 3

$\log \frac{P(X_1 =Yes | X_2 = k)}{P(X_1 =No | X_2 = k)} = u_{1(1)} - u_{1(2)} = 0.2081879 + 0.2081879 =0.4163758$. The interpretation is that the odds of yes over no are exp(0.4163758) = 1.516456.
