---
title: "Stat 502 HW3"
author: "Dongyang Wang"
date: "10/25/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 1

(a)
```{r}
#rm(list = ls())
sleep <- sleep
sleep1 <- sleep[sleep$group == 1,]
sleep2 <- sleep[sleep$group == 2,]
t.test(sleep1$extra, sleep2$extra, var.equal = T)
```
The confidence interval is [-3.363874, 0.203874].

(b)

Since $Y_A, Y_B$ are normally distributed, $Y_A -  Y_B \sim N(\delta, \frac{\sigma^2}{n_A} +\frac{\sigma^2}{n_B})$. To form a standard normal, we have $\frac{Y_A -  Y_B -\delta}{\sqrt{\frac{\sigma^2}{n_A} +\frac{\sigma^2}{n_B}} }$. Since we also know $\frac{n_A +n_B - 2}{\sigma^2} s_P^2$ follows chi-squared distribution with $n_A +n_B - 2$ as the dof, and X and Z are independent. So, by definition of t-statistic, we have $t(Y_A,Y_B) = \frac{\frac{Y_A -  Y_B - \delta}{ \sqrt{\frac{\sigma^2}{n_A} +\frac{\sigma^2}{n_B}} }}{\sqrt{\frac{n_A +n_B - 2}{\sigma^2} s_P^2/(n_A +n_B - 2)}} = \frac{Y_A -  Y_B}{s_p \sqrt{\frac{1}{n_A} +\frac{1}{n_B}}} - \frac{\delta}{s_p \sqrt{\frac{1}{n_A} +\frac{1}{n_B}}}$. As we can observe, this is a non-central t-distribution.

The components are
(i)     Z = $\frac{Y_A -Y_B -\delta}{\sqrt{\frac{\sigma^2}{n_A} +\frac{\sigma^2}{n_B}} }$.
(ii)    X = $\frac{n_A +n_B - 2}{\sigma^2} s_P^2$.
(iii)   The non-centrality parameter is $\gamma = \frac{-\delta}{\sqrt{\frac{\sigma^2}{n_A} +\frac{\sigma^2}{n_B}} }$.

(c) Since power is an increasing function of absolute value of the non centrality parameter $\gamma = |\frac{\delta}{\sqrt{\frac{\sigma^2}{n_A} +\frac{\sigma^2}{n_B}}}|$, we want to maximize $\gamma$. Rewriting with $n_B = N - n_A$, we have $\gamma = |\frac{\delta}{\sigma\sqrt{\frac{1}{n_A} +\frac{1}{N- n_A}}}| = |\frac{\delta}{\sigma\sqrt{\frac{N}{n_A(N- n_A)}}}| = |\frac{\delta \sqrt{n_A(N- n_A)}}{\sigma\sqrt{N}}|$. To maximize this, we let $N-n_A = n_A$, that is $n_A = \frac{1}{2}N$. Hence, we maximized power.

## 3

(a) The model is $y_{ij} = \mu + \tau_{i} + \epsilon_{ij}$. The meaning of $\mu$ means the mean when no treatment is in place. $\tau_{i}$ represent the within treatment variation for the four treatments. Key assumptions include $E(\epsilon_{ij}) = 0$, $Var(\epsilon_{ij}) = \sigma^2$. We also want the sum of $\tau_{i}$ is 0.

(b)
```{r}
A <- c(62,60,63,59,64)
B <- c(65,67,73,65,66)
C <- c(69,66,71,67,67,68,62)
D <- c(66,62,65,61,64,65,63)
effect <- c(A,B,C,D)
diet <- c(rep('A', 5), rep('B', 5),rep('C', 7), rep('D', 7))
data <- data.frame(effect, diet)
plot(as.factor(data$diet), data$effect)
mean(A)
mean(B)
mean(C)
mean(D)
mean(data)
```
There is no big difference between C and B, but among others theire is a difference.

(c)
```{r}
#Group sample variances
sA <- var(A)
sB <- var(B)
sC <- var(C)
sD <- var(D)
sA;sB;sC;sD
# MSE
mse <- (4*sA+4*sB+6*sC+6*sD)/(24 - 4)
mse
```
Group sample variances are 4.3, 11.2, 7.809524, 3.238095. The MSE is 6.414286.

(d)
```{r}
mst <- 5/3*(mean(A) - mean(data$effect))^2 + 5/3*(mean(B) - mean(data$effect))^2
+7/3*(mean(C) - mean(data$effect))^2 + 7/3*(mean(D) - mean(data$effect))^2 
mst
```
MST is 41.90476. It is way larger than MSE, so we can expect to reject the null that the diets are no different.

(e)
```{r}
anova(lm(data$effect ~ data$diet))
```
Yes, since the p-value is 0.002937, we can safely reject the null and say that there is a difference for these four diets.

(f)
Since the fitted values are the mean for each diet, we have
```{r}
data$miu.hat <- c(rep(mean(A), 5), rep(mean(B), 5),rep(mean(C), 7), rep(mean(D), 7))
data$residual <- data$effect - data$miu.hat
qqnorm(data$residual)
```
Yes, the residuals appear to follow a normal distribution, since the points on the plot are pretty much along the diagonal. 

## 4

(a)
```{r}
library(dplyr)
zinc <- readRDS("zinc.RDS")
plot(as.factor(zinc$ZINC), zinc$DIVERSITY)
mean(zinc$DIVERSITY)
# Method 1
meanZ <- zinc %>%
  group_by(ZINC) %>%
  summarize(mean(DIVERSITY))
# 2
aggregate(x= zinc$DIVERSITY,     
         # Specify group indicator
         by = list(zinc$ZINC),      
         # Specify function (i.e. mean)
         FUN = mean)
```
Yes, there appears to be a difference in biodiversity of rivers with different Zinc levels.

(b)
```{r}
#Group sample variances
varZ <- zinc %>%
  group_by(ZINC) %>%
  summarize(var(DIVERSITY))
varZ
#count
numZ <- zinc %>%
  count(ZINC)
#MSE
mse1 <- (0.2354786+0.2081268+0.1980214	+0.2876286			
)/4
mse1
```
Group variances are BACK 0.2354786, HIGH	0.2081268, LOW	0.1980214, MED	0.2876286. MSE is 0.2323139.

(c)
```{r}
meanZ
mst1 <- 8/3*((1.79750 - mean(zinc$DIVERSITY))^2 + (1.28125 - mean(zinc$DIVERSITY))^2 + 
               (2.03250 - mean(zinc$DIVERSITY))^2 + (1.73000 - mean(zinc$DIVERSITY))^2 )
mst1
fratio <- mst1 / mse1
fratio
#verify
anova(lm(zinc$DIVERSITY ~ zinc$ZINC))
```
Yes, I would have expected this F ratio -- it is large such that we can reject the null that there is no difference in biodiversity in zones with different Zinc levels. This is consistent with my observation in the plots and early data exploration of the mean.

(d)

```{r}
zinc.level <- zinc$ZINC
fr <- c()
for(i in 1:1000){
  zinc.sim <- sample(zinc.level)
  fr <- c(fr, anova(lm(zinc$DIVERSITY ~ zinc.sim))$F[1] )
}
mean(fr >= 3.3964)
```
With a p-value of 0.027, we can reject the null and say that Zinc levels do affect biodiversity.
