---
title: "Stat 502 HW4"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 1

(a)
```{r}
zinc <- readRDS("zinc.RDS")
summary(zinc)
n <- 5
m <- 4
N <- m*n
lambda <- n*(1^2 + 5^2 +3^2)/20^2
lambda
alpha <- 0.05
powerF <- 1 - pf(qf(1-alpha, m-1, N-m), m-1,N-m, ncp = lambda)
powerF
```
The power is 0.07123778.

```{r}
powerf <- c()
for (n in 1:200){
  m <- 4
  N < m*n
  lambda <- n*(1^2 + 5^2 +3^2)/20^2
  alpha <- 0.05
  powerf[n] <- 1 - pf(qf(1-alpha, m-1, N-m), m-1,N-m, ncp = lambda)
}
powerf
ind = match(1, powerf >= .8)
ind
```
We need sample size of at least 160.

```{r}
powerf1 <- c()
for (sigma in 1:20){
  m <- 4
  n <- 10
  N < m*n
  lambda <- n*(1^2 + 5^2 +3^2)/sigma^2
  alpha <- 0.05
  powerf1[sigma] <- 1 - pf(qf(1-alpha, m-1, N-m), m-1,N-m, ncp = lambda)
}
powerf1
ind = match(1, powerf1 <= .8) -1
ind
```
Largest value $\sigma$ can take is 5.

## 2

(a)
```{r}
X <- matrix(c(1,1,1,1, 1,1,1,1, 1,1,0,0, 0,0,1,1, 0,0,1,1, 0,0,-1,-1,
              0,0,0,0, 1,1,-1,-1), byrow = F, ncol = 4)
X
```
We have obtained this matrix because of the constraint $\tau_1 - \tau_2 -\tau_3 = \tau_4$. The rows of the matrix are $Y_{ij}$, namely $Y_{11}, Y_{12}$... And the columns are $\mu, \tau_1, \tau_2, \tau_3$.

(b)
```{r}
XT <- t(X)
invxtx <- solve(XT %*% X)
H <- X %*% invxtx %*% XT
H
```
We have obtained the H matrix using $H = X(X^TX)^{-1}X^T$.

(c)
Because of the constraints, we know $\tau_1 = \tau_2 = 2\tau_3 = -10\tau_4$. So we can write everything in terms of $\tau_3$ as follows:
```{r}
X_prime <- matrix(c(1,1,1,1, 1,1,1,1, 2,2,2,2, 1,1,-5,-5), byrow = F, ncol = 2)
X_prime
```

## 3

(a)
```{r}
X_weight <- matrix(c(1,1,1, 0,0,0, 0,0,0, 0,0,0, 1,1,1, 0,0,0, 
                     0,0,0, 0,0,0, 1,1,1), byrow = F, ncol = 3)
X_weight
```

(b)
```{r}
XXTinv <- solve(t(X_weight) %*% X_weight)
Y = matrix(c(78,86.8,103.8,83.7,89,99.2,83.8,81.5,86.2), ncol = 1)
beta <- XXTinv %*% t(X_weight) %*% Y
beta
```
Therefore, the estimate $\hat{\beta_1}$ for $\beta_1$ is the first entry in the above vector, i.e., 89.5333.

(c)
```{r}
mn_1 <- 3*2
sse <- sum((Y[1:3] - mean(c(78,86.8,103.8)))^2 + (Y[4:6] - mean(c(83.7,89,99.2)))^2 
           +  (Y[7:9] - mean(c(83.8,81.5,86.2)))^2)/mn_1
se <- sqrt(sse) * sqrt(XXTinv[1,1])
se
```
The standard error of $\hat{\beta_1}$ is 5.159673.

(d)
By theorem, we know that $\frac{\hat{\beta_1} - \beta_1}{SE(\hat{\beta_1})}$ follows $t_{n-p}$ distribution.
```{r}
n = 9
p = 3
t_c <- qt(0.95,6)
CI = c(beta[1,1] - t_c * se, beta[1,1] + t_c * se)
CI
```
The 90% confidence interval for $\hat{\beta_1}$ is [79.50716, 99.55951].

(e)
Yes they are independent. By our model, we know $\mu_i$ are fixed. Since $\epsilon_{ij}$ are independent, $Y_{ij}$ is also independent because it is a linear function of $\epsilon_{ij}$. For each $\hat{\mu_i}$, we know that $\hat{\mu_i} = X(X^TX)^{-1}Y$ and therefore is a linear function of the Y, which is independent. Therefore, $\hat{\mu}_{beef}$ and $\hat{\mu}_{pork}$ are independent.
