---
title: "hw 7"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 1

### a
This is a Latin square with 2 blocks, store and weekday; 1 factor, special offers A-E.

### b
This is RCBD with 1 block, ground covers, and 1 factor, fertilizer levels. There are 4 observations per combination, and a total of 24 observations.

### c
This is a BIBD with 2 blocks, region and patient; 1 factor ointments A-E.

## 2

### a
We can have a BIBD with 6 blocks, 2 levels in each block. Since N = bk = rg, we have 4 treatments and each treatment appears 3 times.

### b 

Yes, it is a BIBD. First, each tr4eatment appears r=3 times. Second, each treatment appears at most once per block. Third, each pair of treatments appears exactly once without repetition. Also, it satisfies N = bk =rg, where b = 9, k =3, g = 9, r = 3.

```{r}
readRDS("ibd.RDS")
```

## 3

### a
See attachment.

### b
E = ABCD. We want to ensure that main effects and lower order interactions do not get confounded.

### c
See attachment.

### d
ABCDE, where all levels are positive. Added to attachment.

## 4

Since the error is normally distributed, we can construct a t-statistic using the mean between $y_{i2kl}$ and $y_{i1kl}$. Since the variance in the error is known to be 4, we can calculate the variance of each sample mean, which is $\frac{4+4+4+4}{16} = 1$. Therefore, the variance of $y_{i2kl} - y_{i1kl}$ is 2. As a result our test statistic is $\frac{y_{i2kl} - y_{i1kl}}{\sqrt{2}}$ and this follows the standard normal distribution. We can either use a z-test or t-test to conduct the hypothesis testing.

```{r, include= FALSE}
# midterm Q1(3)
y<- c( 26.9,11.4,26.6,23.7,25.3,28.5,14.2,17.9,16.5,21.1,24.3,19.6)
x<-c("B", "A", "B", "A", "B", "B", "B", "A", "A", "A", "B", "A")

sa <- sd(y[x=="A"])
sb <- sd(y[x=="B"])
g31 <- abs(sa^2 -sb^2)
g41 <- abs(sa -sb)


set.seed(1)
g3 <- g4 <- c()
for(nsim in 1:10000)
{
  xsim<-sample(x)
  sa <- sd(y[xsim=="A"])
  sb <- sd(y[xsim=="B"])
  g3[nsim] <- abs(sa^2 -sb^2)
  g4[nsim] <-  abs(sa -sb)
}


mean( abs(g3)>=abs(g31))
mean( abs(g4)>=abs(g41))

meana <- mean(y[x=="A"])
meanb <- mean(y[x=="B"])
g31 <- abs(meana^2 -meanb^2)
g41 <- abs(meana -meanb)

set.seed(1)
g3 <- g4 <- c()

for(nsim in 1:10000)
{
  xsim<-sample(x)
  sa <- mean(y[xsim=="A"])
  sb <- mean(y[xsim=="B"])
  g3[nsim] <- abs(sa^2 -sb^2)
  g4[nsim] <-  abs(sa -sb)
}

mean( abs(g3)>=abs(g31))
mean( abs(g4)>=abs(g41))
```