---
title: "Stat 570 HW6"
author: "Dongyang Wang"
date: "2022-11-12"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Q1

### a

```{r, echo=FALSE}
rm(list = ls())
hodgkins <- c(396, 568, 1212, 171, 554, 1104, 257, 435, 295, 397, 288, 1004, 431,
              795, 1621, 1378, 902, 958, 1283, 2415)
nonhodgkins <- c(375, 375, 752, 208, 151, 116, 736, 192, 315, 1252, 675, 700, 440,
                  771, 688, 426, 410, 979, 377, 503)

res_1 <- rbind(summary(hodgkins),summary(nonhodgkins))
rownames(res_1) <- c("hodgkins", "nonhodgkins")
res_1
```

```{r, echo=FALSE}
boxplot(hodgkins, nonhodgkins, names=c("hodgkins", "nonhodgkins"), main = "Comparison of hodgkins and nonhodgkins distribution")
```
As we can observe, T4 cells are more prevalent in Hodgkins remissions than in others, since it has a higher mean and greater variance. The data is right skewed.

### b

```{r, echo=FALSE}
library(broom)
x <- c(rep(0,length(hodgkins)), rep(1,length(nonhodgkins)))
y <- c(hodgkins, nonhodgkins)

lm1 <- lm(y ~ x)
lm2 <- lm(log(y) ~ x)
lm3 <- lm(sqrt(y) ~ x)
res1_b <- tidy(lm1, conf.int = T, conf.level = 0.90)[2,c(2,6,7)]
res2_b <- tidy(lm2, conf.int = T, conf.level = 0.90)[2,c(2,6,7)]
res3_b <- tidy(lm3, conf.int = T, conf.level = 0.90)[2,c(2,6,7)]

res_b <- rbind(res1_b, res2_b, res3_b)
rownames(res_b) <- c("Original", "Log", "Square Root")
res_b
```
Here we have obtained a table showing the difference between the two categories, by using the beta 1 in the linear model, where I input 0 for the hodgkin's and 1 for nonhodgkin's. In choosing a good mode, we can first consider how we want to interpret the model. That leaves out the third choice because it's hard to interpret. The first one is straightforward and the second one involves the exponential of beta1 for each unit increase in x (i.e, the difference in the two groups, changing from nonhodgkins to hodgkins). Also, considering the right skewness of the data, it's more preferable to use the log transformation for variance stabilization.

### c

To study this we can want to test whether $\beta_1 = 0$, For the Poisson model, we use the canonical log link, i.e., $g(\mu) = \log \mu$. We go back to original scale by $\exp(\beta_0 + \beta_1 x)$ where x= 1 when it is nonhodgkins and 0 for hodgkins. For Gamma, the canonical is reciprocal. $g(\mu) = 1/ \mu$. We go back to original scale by $1/(\beta_0 + \beta_1 x)$ where x= 1 when it is nonhodgkins and 0 for hodgkins. For Inverse Gaussian, the canonical is reciprocal. $g(\mu) = 1/ \mu^2$. We go back to original scale by $(\beta_0 + \beta_1 x)^{-1/2}$ where x= 1 when it is nonhodgkins and 0 for hodgkins.

### d

```{r, echo=FALSE}
model_pois <- glm(y ~ x, family = poisson(link = "log"))
model_gamma <- glm(y ~ x, family = Gamma(link = "inverse"))
model_inv_gauss <- glm(y ~ x, family=inverse.gaussian(link = "1/mu^2"))

model_pois <- c(model_pois$coefficients , model_pois$coefficients + qnorm(0.05)*sqrt(diag(vcov(model_pois))),
  model_pois$coefficients + qnorm(0.95)*sqrt(diag(vcov(model_pois))) )

model_gamma <- c(model_gamma$coefficients , model_gamma$coefficients + qnorm(0.05)*sqrt(diag(vcov(model_gamma))),
  model_gamma$coefficients + qnorm(0.95)*sqrt(diag(vcov(model_gamma))) )

model_inv_gauss <- c(model_inv_gauss$coefficients , model_inv_gauss$coefficients + qnorm(0.05)*sqrt(diag(vcov(model_inv_gauss))),
  model_inv_gauss$coefficients + qnorm(0.95)*sqrt(diag(vcov(model_inv_gauss))) )

res_d <- rbind(model_pois, model_gamma, model_inv_gauss)
colnames(res_d) <- c("b0", "b1", "b0 Lower CI", "b1 Upper CI", "b0 Lower CI", "b1 Upper CI")
res_d
```

Based on the result above, since 0 is not in the interval, we can conclude that the means for two groups are in fact different.

## Q2

### a

Since $\log y_i \sim N(\log(\frac{D}{V} \exp(-k_ex_i)), \sigma^2)$

The likelihood is $L = \Pi_{i=1}^n \frac{1}{\sqrt{2\pi}\sigma} \exp(\frac{(\log y_i - \log D + \log V + k_ex_i)^2}{2\sigma^2})$

The log likelihood is $l = \frac{n}{2}\log \sqrt{2\pi} -n \log \sigma  + \Sigma_{i =1}^n \frac{(\log y_i - \log D + \log V + k_ex_i)^2}{2\sigma^2}$

So, we have the score functions.

$\frac{dl}{dk_e} = -\frac{1}{\sigma^2} \Sigma_{i =1}^n x_i(\log y_i - \log D + \log V + k_ex_i)$

$\frac{dl}{dV} = -\frac{1}{V\sigma^2} \Sigma_{i =1}^n (\log y_i - \log D + \log V + k_ex_i)$

$\frac{dl}{d\sigma^2} =-\frac{n}{2\sigma^2} + \frac{1}{\sigma^4} \Sigma_{i =1}^n (\log y_i - \log D + \log V + k_ex_i)^2$

We also have the information,

$I_{k_ek_e} = - E(\frac{dl}{dk_e dk_e}) = \frac{\Sigma_{i=1}^n x_i^2}{\sigma^2}$

$I_{VV} = - E(\frac{dl}{dV dV}) = - E(\frac{\Sigma_{i =1}^n(\log y_i - \log D + \log V + k_ex_i)}{V^2} - \frac{n}{V^2\sigma^2}) = \frac{n}{V^2\sigma^2}$

$I_{\sigma^2\sigma^2} = - E(\frac{dl}{d\sigma^2 d\sigma^2}) = - E(\frac{n}{2\sigma^4} -\frac{\Sigma_{i =1}^n(\log y_i - \log D + \log V + k_ex_i)}{\sigma^6}) = - E(\frac{n}{2\sigma^4} - \frac{n\sigma^2}{\sigma^6}) = \frac{n}{2\sigma^4}$

$I_{k_eV} =I_{Vk_e} = - E(\frac{dl}{dV dk_e}) = E(-\frac{\Sigma_{i=1}^n x_i}{V\sigma^2}) = \frac{\Sigma_{i=1}^n x_i}{V\sigma^2}$ 

$I_{k_e\sigma^2} =I_{\sigma^2k_e} = - E(\frac{dl}{d\sigma^2 dk_e}) = E(c \Sigma_{i =1}^n(\log y_i - \log D + \log V + k_ex_i)) = 0$ where c is constant.

$I_{V\sigma^2} =I_{\sigma^2V} = - E(\frac{dl}{d\sigma^2 dV}) = E(c \Sigma_{i =1}^n(\log y_i - \log D + \log V + k_ex_i)) = 0$ where c is constant.

### b

By solving the score functions,

$\hat{k_e} = \frac{\Sigma_{i=1}^n \log y_i (x_i - \bar{x_i})}{n\bar{x}^2 - \Sigma_{i=1}^n x_i^2}$

$\hat{V} = \exp (\log D - \frac{1}{n} \Sigma_{i=1}^n \log y_i - \hat{k_e}  \bar{x})$

$\hat{\sigma^2} = \frac{1}{n} \Sigma_{i=1}^n (\log y_i - \log D + \log \hat{V} + \hat{k_e} x_i)^2$

```{r, echo=FALSE}

y <- c(1.63,1.01,0.73,0.55,0.41,0.01,0.06,0.02)
log_y <- log(y)
n <- length(y)
x <- c(2,4,6,8,10,24,28,32)
D <- 30

mle_ke <- sum(log_y*(x-mean(x))) / (n*mean(x)^2-sum(x^2))
mle_V <- exp(log(D)-mean(log_y)-mle_ke*mean(x))
mle_sig2 <- sum((log_y-log(D)+log(mle_V)+mle_ke*x)^2)/n


FIM <- matrix(c(sum(x^2)/mle_sig2, sum(x)/(mle_V*mle_sig2), 0,
                 sum(x)/(mle_V*mle_sig2), n/(mle_V^2*mle_sig2), 0,
                 0, 0, n/(2*mle_sig2^2)),nrow = 3, ncol = 3)

mle_se <- sqrt(diag(solve(FIM)))

res_b <- c(mle_ke, mle_V, mle_sig2) + as.vector(mle_se) %o% c(0,qnorm(0.025),qnorm(0.975))
colnames(res_b) <- c("MLE", "Lower CI", "Upper CI")
rownames(res_b) <- c("Ke", "V", "Sigma^2")
res_b
```

### c

```{r, echo=FALSE}
yhat = exp(log(D/mle_V*exp(-mle_ke*x)))
plot(y ~ x, col="blue", xlab = "Time", ylab = "Drug Concentration")
lines(yhat ~ x, col="red" )

plot(y-yhat ~ x, col="blue", xlab = "Time", ylab = "Residual")
abline(h = 0, col="red")
```


### d

From the plots above, the assumptions are not entirely upheld, since there is one point being outlier, creating a large residual among all points. The other points seem to have good prediction though. The model is reasonable if we do not encounter extreme values.

### e

By invariance of MLE and delta method on FIM, we can obtain the corresponding values of interest.

```{r, echo=FALSE}
Sigma <- solve(FIM[1:2,1:2])
grad_1 <- c(mle_V, mle_ke)
grad_2 <- c(log(2)/mle_ke^2,0)

ci1 <- (mle_V * mle_ke) + sqrt(t(grad_1)%*%Sigma%*%grad_1)[1] %o% c(0,qnorm(0.025),qnorm(0.975))
ci2 <- log(2)/mle_ke + sqrt(t(grad_2)%*%Sigma%*%grad_2)[1] %o% c(0,qnorm(0.025),qnorm(0.975))
res_e <- rbind(ci1, ci2)
res_e
```


## Appendix

### Q1

```{r echo = T, eval=FALSE}
hodgkins <- c(396, 568, 1212, 171, 554, 1104, 257, 435, 295, 397, 288, 1004, 431,
              795, 1621, 1378, 902, 958, 1283, 2415)
nonhodgkins <- c(375, 375, 752, 208, 151, 116, 736, 192, 315, 1252, 675, 700, 440,
                  771, 688, 426, 410, 979, 377, 503)

res_1 <- rbind(summary(hodgkins),summary(nonhodgkins))
rownames(res_1) <- c("hodgkins", "nonhodgkins")
res_1

boxplot(hodgkins, nonhodgkins, names=c("hodgkins", "nonhodgkins"), main = "Comparison of hodgkins and nonhodgkins distribution")

library(broom)
x <- c(rep(0,length(hodgkins)), rep(1,length(nonhodgkins)))
y <- c(hodgkins, nonhodgkins)

lm1 <- lm(y ~ x)
lm2 <- lm(log(y) ~ x)
lm3 <- lm(sqrt(y) ~ x)
res1_b <- tidy(lm1, conf.int = T, conf.level = 0.90)[2,c(2,6,7)]
res2_b <- tidy(lm2, conf.int = T, conf.level = 0.90)[2,c(2,6,7)]
res3_b <- tidy(lm3, conf.int = T, conf.level = 0.90)[2,c(2,6,7)]

res_b <- rbind(res1_b, res2_b, res3_b)
rownames(res_b) <- c("Original", "Log", "Square Root")
res_b

model_pois <- glm(y ~ x, family = poisson(link = "log"))
model_gamma <- glm(y ~ x, family = Gamma(link = "inverse"))
model_inv_gauss <- glm(y ~ x, family=inverse.gaussian(link = "1/mu^2"))

model_pois <- c(model_pois$coefficients , model_pois$coefficients + qnorm(0.05)*sqrt(diag(vcov(model_pois))),
  model_pois$coefficients + qnorm(0.95)*sqrt(diag(vcov(model_pois))) )

model_gamma <- c(model_gamma$coefficients , model_gamma$coefficients + qnorm(0.05)*sqrt(diag(vcov(model_gamma))),
  model_gamma$coefficients + qnorm(0.95)*sqrt(diag(vcov(model_gamma))) )

model_inv_gauss <- c(model_inv_gauss$coefficients , model_inv_gauss$coefficients + qnorm(0.05)*sqrt(diag(vcov(model_inv_gauss))),
  model_inv_gauss$coefficients + qnorm(0.95)*sqrt(diag(vcov(model_inv_gauss))) )

res_d <- rbind(model_pois, model_gamma, model_inv_gauss)
colnames(res_d) <- c("b0", "b1", "b0 Lower CI", "b1 Upper CI", "b0 Lower CI", "b1 Upper CI")
res_d
```

### Q2

```{r echo = T, eval=FALSE}

y <- c(1.63,1.01,0.73,0.55,0.41,0.01,0.06,0.02)
log_y <- log(y)
n <- length(y)
x <- c(2,4,6,8,10,24,28,32)
D <- 30

mle_ke <- sum(log_y*(x-mean(x))) / (n*mean(x)^2-sum(x^2))
mle_V <- exp(log(D)-mean(log_y)-mle_ke*mean(x))
mle_sig2 <- sum((log_y-log(D)+log(mle_V)+mle_ke*x)^2)/n


FIM <- matrix(c(sum(x^2)/mle_sig2, sum(x)/(mle_V*mle_sig2), 0,
                 sum(x)/(mle_V*mle_sig2), n/(mle_V^2*mle_sig2), 0,
                 0, 0, n/(2*mle_sig2^2)),nrow = 3, ncol = 3)

mle_se <- sqrt(diag(solve(FIM)))

res_b <- c(mle_ke, mle_V, mle_sig2) + as.vector(mle_se) %o% c(0,-1.96,1.96)
colnames(res_b) <- c("MLE", "Lower CI", "Upper CI")
rownames(res_b) <- c("Ke", "V", "Sigma^2")
res_b

yhat = exp(log(D/mle_V*exp(-mle_ke*x)))
plot(y ~ x, col="blue", xlab = "Time", ylab = "Drug Concentration")
lines(yhat ~ x, col="red" )

plot(y-yhat ~ x, col="blue", xlab = "Time", ylab = "Residual")
abline(h = 0, col="red")

Sigma <- solve(FIM[1:2,1:2])
grad_1 <- c(mle_V, mle_ke)
grad_2 <- c(log(2)/mle_ke^2,0)

ci1 <- (mle_V * mle_ke) + sqrt(t(grad_1)%*%Sigma%*%grad_1)[1] %o% c(0,qnorm(0.025),qnorm(0.975))
ci2 <- log(2)/mle_ke + sqrt(t(grad_2)%*%Sigma%*%grad_2)[1] %o% c(0,qnorm(0.025),qnorm(0.975))
res_e <- rbind(ci1, ci2)
res_e
```
