---
title: "Stat 536 HW4"
author: "Dongyang Wang"
date: "2022-10-29"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Question 1

```{r}
rm(list = ls())
table1 <- c(28, 656, 28+656, 18, 658, 18+658, 28+18, 656+658, 28+18+656+658)
matrix_d <- matrix(table1,byrow = TRUE, nrow = 3)
colnames(matrix_d) <- c("Yes", "No", "Row Total")
rownames(matrix_d) <- c("Placebo", "Aspirin", "Col Total")
matrix_d
```
Grand total is 1360.

## Question 2

By Frechet lower and upper bounds, 
(1,1) has lower bound 0 and upper bound 46.
(1,2) has lower bound 638 and upper bound 684.
(2,1) has lower bound 0 and upper bound 46.
(2,2) has lower bound 630 and upper bound 676.

## Question 3

Since the smallest range for the cell entries above involves (1,1), the number of T is equal to the number of possible values of (1,1), given current row and column total constraints. In our case, it is 47 because (1,1) can take on any integer between 0 and 46.

The notation with x being the count in (1,1). T = {(x, 684-x, 46 - x, 630+x), x $\in$ {0,1,...46}}.

## Question 4

The following is saturated log linear fit.
```{r}
df = c(28,18,656,658)
df = array(df, c(2,2))
loglin(df,margin=list(c(1,2)),fit=TRUE,param=TRUE)$fit
```


## Question 5

The following is the log linear fit of independence.
```{r}
loglin(df,margin=list(1,2),fit=TRUE,param=TRUE)$fit
```

## Question 6

```{r}
x2 <- 0
for (i in 1:2){
  for (j in 1:2){
    core <- matrix_d[i,3] * matrix_d[3,j]/matrix_d[3,3]
    x2 <- x2 + ((matrix_d[i,j]-core)/sqrt(core))^2
  }
}
pval_6 <- 1 - pchisq(x2, df=1)
pval_6
```
Fail to reject the null that aspirin and myocardial infarction are independent.

## Question 7

```{r}
d <- 0
for (i in 1:2){
  for (j in 1:2){
    core <- matrix_d[i,3] * matrix_d[3,j]/matrix_d[3,3]
    d <- d + 2*matrix_d[i,j] * log(matrix_d[i,j]/core)
  }
}
pval_7 <- 1 - pchisq(d, df=1)
pval_7
```
Fail to reject the null that aspirin and myocardial infarction are independent.

## Question 8

X1: Aspirin Use

X2: Myocardial Infarction

The null: X1 and X2 are independent

The alternative: X1 and X2 are not independent

```{r}
fisher.test(df, simulate.p.value = TRUE, B = 1e5)
```

## Question 9

With the model assuming independence, we have
```{r}
loglin(df,margin=list(c(1,2)),fit=TRUE,param=TRUE)$param$`1`
```
where u1(2) - u1(1) = -0.1096972 -0.1096972 = -0.2193944. This is conditional distribution.

## Question 10

Based on the results from 6,7,8, we know that we fail to reject the null that aspirin and myocardial infarction are independent. That means, the effect of aspirin on the occurrence of myocardial infarction is not significant.

